---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pomcheckr

<!-- badges: start -->
<!-- badges: end -->

The goal of pomcheckr is to mplements the method described in \url{https://stats.idre.ucla.edu/r/dae/ordinal-logistic-regression/} for checking if the proportional odds assumption holds for a cumulative logit model.

## Installation

<!--You can install the released version of pomcheckr from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("pomcheckr")
```
-->

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("melissa-wong/pomcheckr")
```
## Example

This is an example using formula syntax:

```{r example1}
library(pomcheckr)
plot(pomcheck(Species ~ Sepal.Length + Sepal.Width, iris))
```

The following syntax is equivalent:

```{r example2}
plot(pomcheck("Species", c("Sepal.Length", "Sepal.Width"),
              iris))
```

And if you want to examine the `pomcheck` object underlying the plots:

```{r example3}
p <- pomcheck(Species ~ Sepal.Length + Sepal.Width, iris)
summary(p)
p
```

