---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "50%"
)
```

# pomcheckr

<!-- badges: start -->
[![Travis build status](https://travis-ci.com/melissa-wong/pomcheckr.svg?branch=main)](https://travis-ci.com/melissa-wong/pomcheckr)
[![](https://cranlogs.r-pkg.org/badges/grand-total/pomcheckr)](https://cran.rstudio.com/web/packages/pomcheckr/index.html)
<!-- badges: end -->

The goal of pomcheckr is to implement the method described by [UCLA Statistical Consulting](https://stats.idre.ucla.edu/r/dae/ordinal-logistic-regression/) for checking if the proportional odds assumption holds for a cumulative logit model.

## Installation

You can install the released version of pomcheckr from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("pomcheckr")
```

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("melissa-wong/pomcheckr")
```
## Examples

### Default Syntax

The following example uses the default syntax:

```{r example2}
library(pomcheckr)
plot(pomcheck("Species", c("Sepal.Length", "Sepal.Width"),
              iris))
```

### Formula Syntax

This is the equivalent example using formula syntax:

```{r example1}
plot(pomcheck(Species ~ Sepal.Length + Sepal.Width, iris))
```

### Examine Data

This example shows how to access and view the `pomcheck` object underlying the plots:

```{r example3}
p <- pomcheck(Species ~ Sepal.Length + Sepal.Width, iris)
summary(p)
```
```{r}
print(p)
```

